Dark matter web in the Universe

### 2018-06-18 Draft/Chapter1/intro.tex
proof-reading

Thus by the  early 1990s it was clearly demonstrated that the web like structure is a generic type for a wide range of initial conditions in both two-(\citealt{Doroshkevich1989}, \citealt{Melott1990} and \citealt{Beacom1991}) and three- dimensional \citep{Melott1993} cosmological N-body simulations. However it also was demonstrated that the quantitative parameters of the web structures depend on the initial power spectrum. Remarkably the simulations also showed that  adding small scale perturbations does not ruin the large scale structures if the slope of the power spectrum is negative in both two- and three- dimensional simulations.

@@

However, it has to be noted and stressed that almost all the results here pertain to dark matter structures, not baryonic matter distribution of galaxies, stars or gas. While clustering of galaxies is positively correlated with dark matter concentrations, the difference in their dynamics of DM and baryonic gas vast. Perhaps the most important difference in the context of this thesis is the difference of collisional cross-sections. Dark matter particles are collisionless in N-body simulations, hence display multiple velocity streams. The motion of cold baryonic gas resembles that of dark matter only until the first orbit crossing arises. However, streams of gas cannot penetrate through one another, causing layers of neighboring gas layers to collide and result in shock waves. On galactic scales, the gas dynamics is further complicated by processes like radiative cooling and heating, supernova feedback, star formation and more.

>>>>

Thus by the  early 1990s it was clearly demonstrated that the web like structure is a generic type for a wide range of initial conditions in both two-(\citealt{Doroshkevich1980}, \citealt{Doroshkevich1989}, \citealt{Melott1990} and \citealt{Beacom1991}) and three- dimensional \citep{Melott1993} cosmological N-body simulations. However it also was demonstrated that the quantitative parameters of the web structures depend on the initial power spectrum. Remarkably the simulations also showed that  adding small scale perturbations does not ruin the large scale structures if the slope of the power spectrum is negative in both two- and three- dimensional simulations.

@@

However, it has to be noted and stressed that almost all the results here pertain to dark matter structures, not baryonic matter distribution of galaxies, stars or gas. While clustering of galaxies is positively correlated with dark matter concentrations, the difference in the dynamics of DM and baryonic gas vast. Perhaps the most important difference in the context of this thesis is the difference of collisional cross-sections. Dark matter particles are collisionless in N-body simulations, hence display multiple velocity streams. The motion of cold baryonic gas resembles that of dark matter only until the first orbit crossing arises. However, streams of gas cannot penetrate through one another, causing layers of neighboring gas layers to collide and result in shock waves. On galactic scales, the gas dynamics is further complicated by processes like radiative cooling and heating, supernova feedback, star formation and more.

### 2018-06-18 Draft/Chapter5/Paper2017b_Nobmath.tex
proof-reading

(iv) The resulting Hessian eigenvalues characterize the geometry in a four-dimensional
hyper-space of $(-n_{str}, x, y, z)$. The boundary of a region with $\lambda_1 \geq \lambda_2 \geq \lambda_3 > 0$ is a closed convex contour in this hyper-space, and thus it’s projection on to the three-dimensional Lagrangian space is also closed and convex.

>>>>

(iv) The resulting Hessian eigenvalues characterize the geometry in a four-dimensional
hyper-space of $(-n_{str}, x, y, z)$. The boundary of a region with $\lambda_1 \geq \lambda_2 \geq \lambda_3 > 0$ is a closed convex contour in this hyper-space, and thus it's projection on to the three-dimensional Lagrangian space is also closed and convex.

### 2018-06-18 Draft/Chapter6/concl.tex
proof-reading

\begin{figure*} 
\centering\includegraphics[height=15cm]{Chapter2/Plots/caustic_connectivity.png} 
\caption{Caustic surfaces in a small cosmological simulation of $L=20$ Mpc, $N_p = 32^3$. Caustics are filtered by longest side length $l_{max}$ of the triangle consitituting the caustic surface. Top left: all values of $l_{max}$ (all the caustic surfaces), top right: $ l_{max} \leq 1.8  h^{-1} Mpc$, bottom left: $ l_{max} \leq 1.25  h^{-1} Mpc$, bottom right: $ l_{max} \leq 1.0  h^{-1} Mpc$.}
\label{fig:caustic_lmax}
\end{figure*}

>>>>

\begin{figure*} 
\centering\includegraphics[height=15cm]{Chapter2/Plots/caustic_connectivity.png} 
\caption{Caustic surfaces in a small cosmological simulation of $L=20$ Mpc, $N_p = 32^3$. Caustics are filtered by longest side length $l_{max}$ of the triangle consitituting the caustic surface. Top left: all values of $l_{max}$ (all the caustic surfaces), top right: $ l_{max} \leq 1.8  h^{-1} Mpc$, bottom left: $ l_{max} \leq 1.25  h^{-1} Mpc$, bottom right: $ l_{max} \leq 1.0  h^{-1} Mpc$. Multiple colors in bottom right panel indicate isolated inner caustic surfaces, while the other three panels display connected outer cautic structures.}
\label{fig:caustic_lmax}
\end{figure*}

### 2018-06-18 Draft/thesis-ku.tex
proof-reading

My appreciation to the friendly staff and faculty at the Department of Physics and Astronomy, especially Dr. Jennifer Delgado, Kristin Rennells, Kim Hubbel and Joel Saurwein for being kind, supportive and patient with me.

It has been a great pleasure sharing office with Brett, Alex, Yuyu and Keita over the past few years. Stimulating conversations -- physics or otherwise -- was always fun and exciting. Special thanks to Donald and Sinan for the company that kept me sane over the course of PhD.

My time at Lawrence was enjoyable in large part due to many wonderful people I was surrounded by. I've had amazing roommates: Niranjan, Niharika, Darshan and Charu, who

Thank you Mum and Dad, for unconditional love and support throughout graduate school and my life in general.

>>>>

My appreciation to the friendly faculty and staff at the Department of Physics and Astronomy, especially Dr. Jennifer Delgado, Kristin Rennells, Kim Hubbel and Joel Saurwein for being kind and supportive.

My time as a graduate student was enjoyable in large part due to many wonderful people I was surrounded by. Anupama has been an inspiring grad student and a wonderful friend, with whom I have shared moments of frustration, laughter and excitement through the entireity of my PhD. I had a great pleasure sharing offices with Brett, Alex, Yuyu and Keita over the past few years. Stimulating conversations with them -- physics or otherwise -- was always fun and exciting. Special thanks to Donald and Sinan for the company that kept me sane over the course of PhD. I've had amazing roommates: Niranjan, Niharika, Darshan and Charu, who have not only put up with my idiosyncracies, occasionally cooked me dinner and had a great time together in Lawrence.

I very much appreciate the opporitunities provided by Prof. Salman Habib, Prof. Katrin Heitmann and the Cosmology group at Argonne National Laboratories, where I have spent better part of my last year of grad school. I'm really excited about joining this group as a postdoc in the near future.

Thank you Mum and Dad, for unconditional love and support throughout graduate school and my life in general. None of this would have been possible without your care and nurture.
